\documentclass{ximera}

\title{Infinite Dimensional Vector Spaces}

\begin{document}
\begin{abstract}
A short description
\end{abstract}
\maketitle
The process of extending the algebraic and geometrical methods of
linear algebra from matrices to differential or integral operators
consists of going from a finite dimensional vector space, typically
$R^n$, to an infinite dimensional vector space, typically a function
space.

However, a vector space of functions has certain idiosyncrasies
precisely because its dimension is infinite.  These peculiarities are
so important that we must develop the framework in which they arise.

One of the most useful, if not the most useful, framework is the
theory of \emph{Hilbert spaces}, the closest thing to the familiar
finite-dimensional Euclidean spaces. In passing we shall also mention
\emph{metric spaces} and \emph{Banach spaces}.

We shall see that infinite dimensional vector spaces are a powerful
way of organizing the statement and solution of boundary value
problems.  In fact, these spaces are the tool of choice whenever the
linear superposition principle is in control. This happens in signal
processing, in quantum mechanics, electromagnetic wave theory, and
elsewhere.

The most notable peculiarity associated with infinite dimensional vector 
spaces is the issue of \emph{completeness}.

From the viewpoint of physics and engineering, completeness is an
issue of precision in measurements.  We would like to have at our
disposal mathematical concepts which are such that they are capable of
mathematizing natural phenomenona no matter how advanced the technolgy
of measuring their properties.

One of the most useful infinite dimensional vector spaces is Hilbert space.  
To define it, we must have at our disposal the constellation of concepts on 
which it is based.  Let us identify the components of the constellation.

\section{Inner Product Spaces}\label{IPS}\index{inner product space}

An \emph{inner product space} is a vector space, say $\cal H$, together with 
a complex bilinear function $\langle~~,~~\rangle$ having the following 
properties:

\begin{tabbing}
(i)~~~\= $\langle f,g\rangle = \overline{\langle g,f\rangle}$ \=~~\= 
where $f,g \in\cal H$\\
(ii)~~~\> $\langle f,\alpha_1g_1+\alpha_2g_2\rangle = \alpha_1\langle f,g_1
\rangle +\alpha_2\langle f,g_2\rangle$\\
~\> where $\alpha_1$ and $\alpha_2$ are complex numbers\\
(iii)~~~\> $\langle f,f\rangle >0$ if $f\not= \vec 0$\\
~\> and $\langle f,f\rangle =0\Leftrightarrow f=\vec 0$ \quad .
\end{tabbing}



\end{document}